---
title: "Reproducible Research - Peer Assessment 1 - 2nd Attempt"
author: "Urli2016"
date: "January 15, 2016"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

LOADING & PREPROCESSING THE DATA

```{r}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
head(activity)
```

```{r}
names(activity)
```

```{r}
library(lattice)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

What is the mean total number of steps taken per day? Using 2 approaches:
1st approach:
```{r, echo = TRUE}
StepsTotal <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)

hist(StepsTotal$steps, main = "Total Steps Per Day", xlab = "day", col = "blue")
```

MEAN
```{r}
mean(StepsTotal$steps)
```

MEDIAN
```{r}
median(StepsTotal$steps)
```

Result mean [1] 10766
Result median [1] 10765

2nd approach: creating a data frame w/weekdays and total steps per day:
```{r}
steps <- rep(NA, 61)
day <- rep("NA", 61)
stepsday <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
length(stepsday)
```

Result: [1] 61

```{r}
for (i in 1:61) {
    steps[i] <- stepsday[[i]]
    day[i] <- names(stepsday)[i]
}
```

data frame:
```{r}
df <- data.frame(day, steps)
head(df)
```

         day steps
1 2012-10-01     0
2 2012-10-02   126
3 2012-10-03 11352
4 2012-10-04 12116
5 2012-10-05 13294
6 2012-10-06 15420

HISTOGRAM (little bit different because of some missing data)
```{r, echo = TRUE}
hist(df$steps, main = "Total steps by day", xlab = "day", col = "green")
```


AVERAGE DAILY ACTIVITY PATTERN:
MEAN OF STEPS AND TIME SERIES PLOT
```{r}
time_series <- tapply(activity$steps, activity$internal, mean, na.rm = TRUE)
```

```{r, echo = TRUE}
plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", 
    ylab = "Average across all Days", main = "Average number of steps taken", 
    col = "red")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- which.max(time_series)
names(max_interval)
```

Result [1] 2304


STRATEGY FOR FILLING IN ALL THE MISSING VALUES IN THE DATA SET
!!First Na replaced by mean in 5 min interval!!
```{r}
StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity)) {
    obs <- activity[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```

CREATING A NEW DATA SET EQUAL TO THE ORIGINAL ONE CONTAINING THE MISSING VALUES
```{r}
new_activity <- activity
new_activity$steps <- fillNA
```

IMPACT OF IMPUTING MISSING DATA
```{r}
StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
```

RELATED HISTOGRAM
```{r, echo = TRUE}
hist(StepsTotal2$steps, main = "Total Steps Per Day", xlab = "day", col = "magenta")
```

MEAN
```{r}
mean(StepsTotal2$steps)
```

MEDIAN
```{r}
median(StepsTotal2$steps)
```

Result mean [1] 10766
Result median [1] 10766

The median has changed a bit, both results (mean and median) are now the same.


DIFFERENCES IN ACTIVITY PATTERNS BETWEEN WEEKDAYS AND WEEKENDS
```{r}
day <- weekdays(activity$date)
daylevel <- vector()
for (i in 1:nrow(activity)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```


RELATED PANEL BLOT containing a time series plot of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r, echo = TRUE}
plot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```
